---
title: "Example Project"
author: "Sister Johnson"
date: "4/27/2021"
output: 
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r class.source = 'fold-hide'}
#This code was run previously to install the data4led package.
  #devtools::install_github('byuidatascience/data4led')
# We only need to install the package once. After that we just call the package using the library() function in R.

library(data4led)
dist <- led_time(2100)
```

We have measurements for 202 light bulbs measured 2104 hours after these light bulbs were turned on. It may be interesting to note that 2104 hours is almost 3 months. The intensity of the bulb as a percent of its original brightness is recorded.

We know that there is variation and uncertainty in measurements. We visualize the variation in the measurements of these 202 light bulbs using a density histogram shown below.

```{r}
hist(dist$percent_intensity,probability = TRUE)
```

We see the light bulbs have intensities ranging from 99.5% to 103% of their original intensities with most of the light bulbs having intensity between 101% and 102%. And there are very few light bulbs with intensities less than 100% or more than 103%. This means most of the light bulbs are a light bulbs are a little brighter after 2014 hours of use compared to when they were first turned on.
